---
title: "Analysis: framing demo"
author: "Your name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(jsonlite)
```

This notebook reads the CSVs written by `npm run export -- --experiment framing-demo`
and runs the planned analyses for the demo experiment. Replace the sections below with
the confirmatory analysis you preregistered; keep exploratory analyses clearly separated.

## Load data

`trials.csv` is long format: one row per jsPsych trial, one participant per `participant_id`.
Columns that hold objects (for example the survey plugin's `response`) are JSON strings.

```{r load}
experiment_id <- "framing-demo"
raw_dir <- file.path("..", "data", "raw", experiment_id)

participants <- read_csv(file.path(raw_dir, "participants.csv"), show_col_types = FALSE)
trials <- read_csv(file.path(raw_dir, "trials.csv"), show_col_types = FALSE)
errors <- read_csv(file.path(raw_dir, "errors.csv"), show_col_types = FALSE)

glimpse(participants)
```

## Exclusions

Document every exclusion. Record how many participants each rule removes.

```{r exclusions}
n_started <- nrow(participants)
complete <- participants |> filter(completed)
n_complete <- nrow(complete)

# Example attention/performance criterion: at least 6 of 8 lexical decision trials correct.
ld_accuracy <- trials |>
  filter(task == "lexical_decision") |>
  group_by(participant_id) |>
  summarise(ld_acc = mean(correct == TRUE | correct == "true", na.rm = TRUE), .groups = "drop")

keep <- complete |>
  left_join(ld_accuracy, by = "participant_id") |>
  filter(ld_acc >= 0.75)

# Report the funnel by condition: differential attrition or exclusion is a threat to validity.
bind_rows(
  participants |> mutate(step = "started"),
  complete |> mutate(step = "completed"),
  keep |> mutate(step = "passed accuracy criterion")
) |>
  count(step, condition) |>
  pivot_wider(names_from = condition, values_from = n, values_fill = 0)
```

## Demographics

```{r demographics}
demo <- trials |>
  filter(task == "demographics", participant_id %in% keep$participant_id) |>
  mutate(resp = map(response, ~ fromJSON(.x))) |>
  transmute(
    participant_id,
    age = map_dbl(resp, ~ as.numeric(.x$age %||% NA)),
    gender = map_chr(resp, ~ as.character(.x$gender %||% NA)),
    native_english = map_lgl(resp, ~ isTRUE(.x$native_english))
  )

demo |> summarise(n = n(), mean_age = mean(age, na.rm = TRUE), sd_age = sd(age, na.rm = TRUE))
demo |> count(gender)
```

## Confirmatory analysis: framing effect

Key test (preregistered): proportion choosing the risky program (B) is higher in the loss
frame than in the gain frame. Chi-square test of independence on the 2 x 2 table.

```{r framing}
framing <- trials |>
  filter(task == "framing", participant_id %in% keep$participant_id) |>
  select(participant_id, condition, choice)

framing |> count(condition, choice) |>
  group_by(condition) |> mutate(prop = n / sum(n))

tab <- table(framing$condition, framing$choice)
tab
if (all(dim(tab) == c(2, 2))) chisq.test(tab)

# Effect size with a confidence interval: the difference in the proportion choosing the risky
# program (loss minus gain). Compare this interval to the original effect, not just the p value.
risky <- framing |> group_by(condition) |> summarise(risky = sum(choice == "risky"), n = n(), .groups = "drop")
if (nrow(risky) == 2 && all(risky$n > 0)) {
  pt <- prop.test(x = risky$risky[risky$condition == c("loss", "gain")],
                  n = risky$n[risky$condition == c("loss", "gain")], correct = FALSE)
  tibble(estimate_loss_minus_gain = diff(rev(pt$estimate)), ci_low = pt$conf.int[1], ci_high = pt$conf.int[2])
}

ggplot(framing |> count(condition, choice) |> group_by(condition) |> mutate(prop = n / sum(n)),
       aes(x = condition, y = prop, fill = choice)) +
  geom_col() +
  labs(x = "Frame", y = "Proportion of choices", fill = "Choice") +
  theme_minimal()
```

## Exploratory: lexical decision RTs

```{r rt}
ld <- trials |>
  filter(task == "lexical_decision", participant_id %in% keep$participant_id) |>
  mutate(correct = as.logical(correct), is_word = as.logical(is_word), rt = as.numeric(rt))

ld |>
  filter(correct, !is.na(rt)) |>
  group_by(is_word) |>
  summarise(mean_rt = mean(rt), sd_rt = sd(rt), n = n())

ggplot(ld |> filter(correct, !is.na(rt)), aes(x = rt, fill = is_word)) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  labs(x = "RT (ms)", fill = "Real word") +
  theme_minimal()
```

## Data quality

Client-side errors reported by the experiment page, if any.

```{r errors}
errors |> select(any_of(c("at", "uid", "trial_index", "message"))) |> head(20)
```

## Session info

```{r session}
sessionInfo()
```
